Report and Figures

### 2016-07-01 tex/results.tex
2nd paper draft

@@

	We assembled block-group level data from the 2008-2012 American Community Survey (ACS), conducted by the U.S. Census Bureau, on race and ethnicity for counties corresponding to 28 large US cities. At this early stage, the counties used were chosen to correspond to the most densely populated urban cores; in future developments we could standardize to conduct analyses for Metropolitan Statistical Areas. We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'. For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$. To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 1km. We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population. A technical specification of this approach is provided in Appendix 1, and a complete summary of cities and their associated information measures is provided in Appendix 3.

		\begin{figure}
			\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
			\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
			\label{fig:info_cross}
		\end{figure}	
	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$. An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.  On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$. For example, the cities of Detroit and Philadelphia provide a striking contrast. While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher. This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure. These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities.

	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density. In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities. A clear trend is evident, with $J(X,Y)$ growing linearly with the logarithm of density. We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix. One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own.

@@

\subsection{Complexity and Neighborhood Identification}

	Figure \ref{fig:cluster_map} shows an illustrative clustering into ten regions according to equation \eqref{eq:info_dist} for the city of Chicago.

		\begin{figure}
			\centering
			\includegraphics[width=\textwidth]{figs/example_cluster_map.png}
			\caption{The regions determined by our hierarchical clustering algorithm reflect high-level racial trends across space. The clusters on the right sharply segment predominantly Black, Hispanic, and White neighborhoods in Central Chicago.}
			\label{fig:cluster_map}
		\end{figure}

		We expect that spatially complex cities with high $J(X,Y)$ would require more complicated models in order to convey similar amounts of demographic information. Testing this expectation requires a quality measure defined on cluster models for each city. The natural measure of loss for a clustering is the mutual information $I(C,Y)$ between cluster and racial labels. Cities vary sharply in the level of cluster detail needed to convey equal amounts of information about race, as shown in figure \ref{fig:clusters}. Each clustering has approximately equal $I(C,Y)$. In sharply segregated Atlanta, just two clusters--one predominantly white, the other predominantly black--suffice to convey more information than do five in Boston, where the clusters are substantially more nuanced. Cluster 1 is predominantly white, cluster 4 majority Hispanic, and cluster 5 overwhelmingly black. Cluster 2 is again majority white, but is unique in that its minority citizens are almost all Hispanic, while cluster 3 is approximately equally black and Hispanic.

		These clusterings can also reveal phenomena related to traditional concepts in segregation studies, such as affinity and concentration. In Figure \ref{fig:clusters}, across all cities, Asians are much more likely to be found in predominantly white clusters than they are in predominantly black or Hispanic ones, indicating significant spatial integration between these two groups. We can observe the phenomenon of concentration -- in which marginalized groups tend to be more highly concentrated in their regions than majority groups are in theirs -- in Atlanta and Detroit. In Atlanta, whites make up 61\% of cluster 1, while blacks make up 83\% of cluster 2. Similarly in Detroit, whites make up 70\% of cluster 1, while blacks make up a full 86\% of cluster 2.

@@

@@

		\begin{figure}
			\centering
			\includegraphics[width=\textwidth]{figs/info_performance_adjed.png}
			\input{figs/adjusted_regression.txt}

@@

>>>>

\nocite{Bivand2014b,Bivand2014a,Bivand2014} % need to cite the census in here too

\subsection{Data Used}
	We assembled block-group level data from the 2010-2014 American Community Survey (ACS), conducted by the U.S. Census Bureau, on race and ethnicity for counties housing 51 large US cities. We then aggregated the detailed racial and ethnic groups into five meta-categories: `Asian', `Black', `Hispanic', `Other', and `White'.

@@

	 For each city, we computed the entropy $H(Y)$ and the mutual information $I(X,Y)$. To compute the estimated aggregate Fisher information $J(X,Y)$, we tiled the map with a hexagonal grid of cell radius 0.5km. We then computed the estimated mutual information within each grid cell, and averaged the results weighted by population population. A technical specification of this approach is provided in Appendix 1, and an illustration of it in Figure \ref{fig:method}.

	 \begin{figure}
		\centering
		\includegraphics[width=.8\textwidth]{figs/method_illustration.png}
		\caption{Illustration of methodology for estimating local information in Suffolk County, MA (Boston). We first cover the map in a hexgrid, and compute the mutual information within each hex from the Census tracts that overlap the hex, weighted by density. }
		\label{fig:method}
	\end{figure}

	Figure \ref{fig:info_cross} shows the relationship of the global spatial variability $I(X,Y)$ and mean local variability $J(X,Y)$. An overall positive trend is evident, reflecting the fact that global spatial variability is a prerequisite for local variability: if the city is uniform (like Figure \ref{fig:toy}(a)), then no local differences either exist.  On the other hand, there are substantial variations in $J(X,Y)$ even in cities with comparable global variability $I(X,Y)$. For example, the cities of Detroit and Philadelphia provide a striking contrast. While they have comparable global variability $I(X,Y)$, Philadelphia's $J(X,Y)$ is substantially higher. This reflects the fact that Detroit is composed of large, highly-segregated, monoracial neighborhoods, whereas Philadelphia has a much more fine-grained, intricate neighborhood structure. These patterns illustrate how combinations of the information measures $H(Y)$, $I(X,Y)$, and $J(X,Y)$ can be used to construct taxonomies diversity for American cities. 
	\begin{figure}
		\includegraphics[width=1\textwidth]{figs/mutual_fisher.png}
		\caption{Relationship of global mutual information $I(X,Y)$ and mean local information $J(X,Y)$.} 
		\label{fig:info_cross}
	\end{figure}

@@

	Intriguingly, the mean local variability $J(X,Y)$ appears obey a scaling relation with respect to urban density. In Figure \ref{fig:density}, we plot $J(X,Y)$ against the population density $\rho$ of our studied cities. It appears that $J(X,Y)$ grows linearly with the logarithm of density. We interpret this trend as reflecting a compression of social space in dense urban areas: the same structure of racial variability fits into much less geographical area in New York than in Phoenix. One aspect of diversity in large, dense cities is that one need walk much less distance in order to reach a neighborhood with substantially different racial trends than one's own. It may be of interest for later studies to consider why this trend might arise through dynamical processes of urban formation.

\subsection{Learning From Natural Neighborhoods}

	Figure \ref{fig:cluster_map} shows an illustrative clustering of Wayne County (housing the city of Detroit) into six regions using the greedy information maximization procedure defined by equation \eqref{eq:info_dist}. The procedure cleanly divides the city into racially coherent zones. Cluster A is predominantly white and B predominantly black, demarcating the major racial divide in the city. Cluster C is a small transitional zone in which the two races overlap. Clusters D and E are the demographically distinct independent cities of Hamtramck and Grosse Pointe, respectively, while Cluster F is a Hispanic community known as Mexicantown. The existence of clusters like C reflects that not every racially coherent tract is interpretable as a ``neighborhood''; some may be more naturally thought as transitions between neighborhoods.

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{figs/example_cluster_map.png}
		\caption{Clustering of Detroit into 6 neighborhoods based on race.}
		\label{fig:cluster_map}
	\end{figure}

	Cities vary sharply in the level of cluster detail needed to convey equal amounts of demographic information. Figure \ref{fig:clusters} shows example cluster divisions for four major cities. Each clustering conveys approximately equal demographic information, as measured by the mutual information $I(C,Y)$ between cluster labels and racial distribution. In sharply segregated Atlanta, just two clusters--one predominantly white, the other predominantly black--suffice to convey as much information as five in Boston, where the clusters are substantially more nuanced. Boston's Cluster A is predominantly white, cluster D majority Hispanic, and cluster E majority black. Cluster B is again majority white, but is distinct in that its minority citizens are almost all Hispanic. Cluster C is approximately equally black, Hispanic, and white.

	Clusterings such as those shown in Figure \ref{fig:clusters} shed useful light on traditional concerns in the study of segregation. We have already argued that $I(X,Y)$ and $J(X,Y)$ naturally encode the high-level ideas of evenness and exposure endorsed by \cite{Reardon2004}. We now consider the further concepts of clustering and concentration. According to \cite{Press2009a}, ``\emph{clustering} measures the degree to which minority areas are located adjacent to one another'' and ``\emph{concentration} refers to the degree of a group's agglomeration in urban space'' (pages 309-310). When areas with similar racial composition are located close to one another, fewer spatial clusters are necessary to convey equivalent information about racial trends. Thus, \ref{fig:clusters} suggests that Atlanta is more clustered than Boston, a suggestion that we can quantify using the AUC methodology developed below. The concept of concentration invites us to consider the composition and density of each cluster. In Detroit, for example, whites make up 70\% of ``their'' Cluster A, while black residents make up a full 86\% of ``their'' Cluster B. Cluster B is also more concentrated in that it is more densely populated than Cluster A: it contains 34\% of the Detroit's population, but accounts for just 19\% of land in the analyzed area. Thus, in Detroit, ``black'' neighborhoods tend to be more racially homogeneous and more densely packed in urban space than ``white'' ones. Finally, we note that these clusterings allow us to examine the somewhat abstract concept of ``exposure'' in considerable detail. Figure \ref{fig:clusters} indicates that across all cities, Asians are much more likely to be found in predominantly white clusters than they are in predominantly black or Hispanic ones, indicating significant spatial integration between these two groups. In Chicago, some groups of Hispanics are exposed primarily to Hispanics and other whites (Cluster B), while other groups are exposed primarily to black residents (Cluster D).

@@

\subsection{Complexity and Neighborhood Structure}

	We expect that spatially complex cities with high $J(X,Y)$ would require more complicated models in order to capture similar levels of spatiosocial structure. Testing this expectation requires a quality measure defined on cluster models for each city. The natural measure of loss for a clustering is the mutual information $I(C,Y)$ between cluster and racial labels.

@@

	\begin{figure}
		\centering
		% \includegraphics[width=\textwidth]{figs/info_performance_adjed.png}
		\input{figs/adjusted_regression.txt}

@@

	Add in a note here about the applicability of this method. Generally, it makes sense on data sets where the resolution is high but gratuitous, allowing us to aggregate up in a non-arbitrary way. Thus, we need the census data to have greater resolution than the ``true'' scale of urban variability.
